---
title: "Hotels and neighbourbood"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
library(ggplot2)


hf100 <- read.csv("~/Downloads/hotel_food_100.csv")
hf <- na.omit(hf100)

ggplot(hf, aes(food_price)) +
  geom_histogram(binwidth = 1, color ="blue", fill = "lightblue") +
  facet_wrap(~hotel_price) +
  ggtitle("Price distribution of hotels and neighbourhood restaurants") +
  xlab("Dollar sign of restaurants") + 
  ylab("Counts of hotels within 100^2pi sq meters") +
  stat_bin(aes(y=..count.., label=..count..), geom="text", vjust=-.5, binwidth = 1) 


ggplot(hf, aes(food_rating)) +
  geom_histogram(binwidth = 1, color ="blue", fill = "lightblue") +
  facet_wrap(~hotel_rating2, nrow = 2) +
  ggtitle("Rating distribution of hotels and neighbourhood restaurants") +
  xlab("Rating level of restaurants") + 
  ylab("Counts of hotels within 100^2pi sq meters") +
  stat_bin(aes(y=..count.., label=..count..), geom="text", vjust=-.5, binwidth = 1) 

ggplot(hf, aes(x = food_price,y = food_review_count)) +
  geom_col(fill = "lightblue") +
  facet_wrap(~hotel_rating2, nrow = 2) +
  ggtitle("Review Counts of hotels and neighbourhood restaurants") +
  xlab("Review Counts of restaurants") + 
  ylab("Counts of hotels within 100^2pi sq meters") 
```

We could find some interesting facts here. 
1) The higher the rating of the restaurant it is, it is not necessary that the ratings of the neighbourhood restaurants are higher.
Taking the rating level 5 as an example, there are not that many reviews nor high ratings. One of the reasons might be not that many restaurants rated as 5 or 1. A lot of the hotels are rated as level 3.
2) There are a lot of restaurants rated as level 3, and the majority of neighbourhood restaurants around level 3 hotels are also on level 3.
3) The restaurants in Time Square are not too expensive. The most expensive ones are with three dollar signs. 
4. The distributions of each price level of hotels are very similar. 4 dollar sign restaurants are very rare, and the proportion of 2 dollar sign restaurants is nearly 50% of the data at each level. 
5) Hotels with 2 dollar signs have the most counts in Time Square. 
